---
title: "Fest verankert – Eine Simulationsstudie zu den Voraussetzungen für robustes Test-Equating"
author: "Dominik Weber"
date: "März 2021"
output: html_document
---

# Loading packages

```{r message = FALSE, warning = FALSE}
library(car)
library(catIrt)
library(dplyr)
library(effectsize)
library(equateIRT)
library(ggplot2)
library(jmv)
library(ltm)
library(MOTE)
library(psych)
library(tidyr)
```

```{r include = FALSE}
# Preparing structure of the equating design

cleanData = function(sample){
    return(sample[, colSums(is.na(sample)) == 0])
}
directGetDiff = function(subsamples){
  estc = list()
  estv = list()
  # subsamples = lapply(subsamples, cleanData)
  for (i in 1:length(subsamples)) {
    subsample = subsamples[[i]]
    est = import.ltm(rasch(subsample[, 1:(ncol(subsample)-2)]), display = FALSE)
    estc[[i]] = est$coef
    estv[[i]] = est$var
  }
  test <- paste("test", 1:k, sep = "")
  links = character(length(subsamples) - 1)
  for (i in 1:length(subsamples) - 1) {
    links[i] = paste(test[i + 1], test[1], sep = ".")
  }
  
  mod1pl <- modIRT(coef = estc, var = estv, names = test, display = FALSE)
  direclist1pl <- alldirec(mods = mod1pl, method = "mean-mean")
  
  result = data.frame(Item = 1:200, Diff = 1:200)
  for (i in 1:length(links)) {
    itmData = itm(direclist1pl, links[i])
    itmData = itmData[-grep("Dscrmn", itmData$Item),]
    if(i == 1){
      result[1:20, 1] = itmData[!is.na(itmData[3]), 1]
      result[1:20, 2] = itmData[!is.na(itmData[3]), 3]
    }
    row = (i) * 20 + 1
    result[row:(row+19), 1] = itmData[is.na(itmData[3]), 1]
    result[row:(row+19), 2] = itmData[is.na(itmData[3]), 4]
  }
  result[, 1] = as.numeric(gsub(pattern = "Dffclt.Item", replacement = "", x = result[, 1]))
  # attach(result)
  result = result[order(result$Item),]
  # detach(result)
  return(result)
}
indirectGetDiff = function(subsamples, items = 25) {
    estc = list()
    estv = list()
    # subsamples = lapply(subsamples, cleanData)
    for (i in 1:length(subsamples)) {
        subsample = subsamples[[i]]
        est = import.ltm(rasch(subsample[, 1:(ncol(subsample) - 2)]), display = FALSE)
        estc[[i]] = est$coef
        estv[[i]] = est$var
    }
    rm(subsample, est)
    test <- paste("test", 1:10, sep = "")
    links = character(length(subsamples) - 1)
    for (i in 1:length(subsamples) - 1) {
        links[i] = paste(test[i + 1], test[1], sep = ".")
    }
    mod1pl <- modIRT(coef = estc, var = estv, names = test, display = FALSE)
    rm(estc, estv)
    direclist1pl <- alldirec(mods = mod1pl, method = "mean-mean")
    
    cecs = list()
    for (i in 3:10) {
        cecs[[i-2]] = chainec(r = i, direclist = direclist1pl, f2 = "test1", f1 = test[i])
    }
    paths=list()
    paths[[1]] = "test3.test2.test1"
    for (i in 2:8) {
        paths[[i]] = paste(test[i + 2],paths[[i - 1]], sep = ".")
    }
    
    result = data.frame(Item = 1:200, Diff = 1:200)
    
    itmData = itm(direclist1pl,"test2.test1")
    itmData = itmData[-grep("Dscrmn", itmData$Item),]
    result[1:items, 1] = itmData[!is.na(itmData[3]), 1]
    result[1:items, 2] = itmData[!is.na(itmData[3]), 3]
    result[(items + 1):(items + 20), 1] = itmData[is.na(itmData[3]), 1]
    result[(items + 1):(items + 20), 2] = itmData[is.na(itmData[3]), 4]
    
    pathsLen = length(paths)
    for (i in 1:length(paths)) {
        if(i < pathsLen){
            row = (items + 21) + items * (i - 1)
            itmData = itm(cecs[[i]],paths[[i]], bistype = "weighted")
            itmData = itmData[-grep("Dscrmn", itmData$Item),]
            result[row:(row + items - 1), 1] = itmData[is.na(itmData[3]), 1]
            result[row:(row + items - 1), 2] = itmData[is.na(itmData[3]), 4]
        }
        else if(i == pathsLen){
            row = (items + 21) + items * (i - 1)
            itmData = itm(cecs[[i]],paths[[i]], bistype = "weighted")
            itmData = itmData[-grep("Dscrmn", itmData$Item),]
            result[row:(row + 19), 1] = itmData[is.na(itmData[3]), 1]
            result[row:(row + 19), 2] = itmData[is.na(itmData[3]), 4]
        }
    }
    
    result[, 1] = as.numeric(gsub(pattern = "Dffclt.Item", replacement = "", x = result[, 1]))
    result = result[order(result$Item),]
    result <- result %>% group_by(Item) %>% summarise_each(list(mean = mean))
}
```


# Settings

## Global parameters

In order to simulate data certain parameters need to be entered. The software must know how many cases (N) and how many samples (k) are to be generated. These variables can be be set freely with values bigger than zero. Also, in order to gain reproducible results the randomization seed in R can be fixed.

```{r}
N = 50
k = 10
set.seed(123456)
```

## IRT specific parameters

Depending on which IRT model you want to simulate data for, different parameters are required.
For the 1PL (Rasch) model, item discrimination (a) and guessing parameters (c) are assumed to be constant and are usually fixed to 1 and 0 respectively. Only the item difficulty (b) varies across items and needs to be defined.

```{r}
a = 1
b = c(rep(-1.34, 20), rep(-0.36, 40), rep(-0.13, 80), rep(1.06, 40), rep(1.87, 20))
c = 0
params = cbind(a, b, c)
rm(a, b, c)
```

## Creating item sets

Since different versions of test equating are to be examined, a list of item sets needs to be defined.

```{r include = FALSE}
directSet1 = list(
  c(1:2,  21:24,61:68,  141:144,181:182),
  c(3:4,  25:28,69:76,  145:148,183:184,1,21,61,141,181),
  c(5:6,  29:32,77:84,  149:152,185:186,1,21,61,141,181),
  c(7:8,  33:36,85:92,  153:156,187:188,1,21,61,141,181),
  c(9:10, 37:40,93:100, 157:160,189:190,1,21,61,141,181),
  c(11:12,41:44,101:108,161:164,191:192,1,21,61,141,181),
  c(13:14,45:48,109:116,165:168,193:194,1,21,61,141,181),
  c(15:16,49:52,117:124,169:172,195:196,1,21,61,141,181),
  c(17:18,53:56,125:132,173:176,197:198,1,21,61,141,181),
  c(19:20,57:60,133:140,177:180,199:200,1,21,61,141,181)
)
directSet2 = list(
  c(1:2,  21:24,61:68,  141:144,181:182),
  c(3:4,  25:28,69:76,  145:148,183:184,1,21,141,181),
  c(5:6,  29:32,77:84,  149:152,185:186,1,21,141,181),
  c(7:8,  33:36,85:92,  153:156,187:188,1,21,141,181),
  c(9:10, 37:40,93:100, 157:160,189:190,1,21,141,181),
  c(11:12,41:44,101:108,161:164,191:192,1,21,141,181),
  c(13:14,45:48,109:116,165:168,193:194,1,21,141,181),
  c(15:16,49:52,117:124,169:172,195:196,1,21,141,181),
  c(17:18,53:56,125:132,173:176,197:198,1,21,141,181),
  c(19:20,57:60,133:140,177:180,199:200,1,21,141,181)
)
directSet3 = list(
  c(1:2,  21:24,61:68,  141:144,181:182),
  c(3:4,  25:28,69:76,  145:148,183:184,21,61,141),
  c(5:6,  29:32,77:84,  149:152,185:186,21,61,141),
  c(7:8,  33:36,85:92,  153:156,187:188,21,61,141),
  c(9:10, 37:40,93:100, 157:160,189:190,21,61,141),
  c(11:12,41:44,101:108,161:164,191:192,21,61,141),
  c(13:14,45:48,109:116,165:168,193:194,21,61,141),
  c(15:16,49:52,117:124,169:172,195:196,21,61,141),
  c(17:18,53:56,125:132,173:176,197:198,21,61,141),
  c(19:20,57:60,133:140,177:180,199:200,21,61,141)
)
directSet4 = list(
  c(1:2,  21:24,61:68,  141:144,181:182),
  c(3:4,  25:28,69:76,  145:148,183:184,1,181),
  c(5:6,  29:32,77:84,  149:152,185:186,1,181),
  c(7:8,  33:36,85:92,  153:156,187:188,1,181),
  c(9:10, 37:40,93:100, 157:160,189:190,1,181),
  c(11:12,41:44,101:108,161:164,191:192,1,181),
  c(13:14,45:48,109:116,165:168,193:194,1,181),
  c(15:16,49:52,117:124,169:172,195:196,1,181),
  c(17:18,53:56,125:132,173:176,197:198,1,181),
  c(19:20,57:60,133:140,177:180,199:200,1,181)
)
directSet5 = list(
  c(1:2,  21:24,61:68,  141:144,181:182),
  c(3:4,  25:28,69:76,  145:148,183:184,61),
  c(5:6,  29:32,77:84,  149:152,185:186,61),
  c(7:8,  33:36,85:92,  153:156,187:188,61),
  c(9:10, 37:40,93:100, 157:160,189:190,61),
  c(11:12,41:44,101:108,161:164,191:192,61),
  c(13:14,45:48,109:116,165:168,193:194,61),
  c(15:16,49:52,117:124,169:172,195:196,61),
  c(17:18,53:56,125:132,173:176,197:198,61),
  c(19:20,57:60,133:140,177:180,199:200,61)
)
indirectSet1 = list(
  c(1:2,  21:24,61:68,  141:144,181:182,20,60,140,180,200),
  c(3:4,  25:28,69:76,  145:148,183:184,2, 24,68, 144,182),
  c(5:6,  29:32,77:84,  149:152,185:186,4, 28,76, 148,184),
  c(7:8,  33:36,85:92,  153:156,187:188,6, 32,84, 152,186),
  c(9:10, 37:40,93:100, 157:160,189:190,8, 36,92, 156,188),
  c(11:12,41:44,101:108,161:164,191:192,10,40,100,160,190),
  c(13:14,45:48,109:116,165:168,193:194,12,44,108,164,192),
  c(15:16,49:52,117:124,169:172,195:196,14,48,116,168,194),
  c(17:18,53:56,125:132,173:176,197:198,16,52,124,172,196),
  c(19:20,57:60,133:140,177:180,199:200,18,56,132,176,198)
)
indirectSet2 = list(
  c(1:2,  21:24,61:68,  141:144,181:182,20,60,180,200),
  c(3:4,  25:28,69:76,  145:148,183:184,2, 24,144,182),
  c(5:6,  29:32,77:84,  149:152,185:186,4, 28,148,184),
  c(7:8,  33:36,85:92,  153:156,187:188,6, 32,152,186),
  c(9:10, 37:40,93:100, 157:160,189:190,8, 36,156,188),
  c(11:12,41:44,101:108,161:164,191:192,10,40,160,190),
  c(13:14,45:48,109:116,165:168,193:194,12,44,164,192),
  c(15:16,49:52,117:124,169:172,195:196,14,48,168,194),
  c(17:18,53:56,125:132,173:176,197:198,16,52,172,196),
  c(19:20,57:60,133:140,177:180,199:200,18,56,176,198)
)
indirectSet3 = list(
  c(1:2,  21:24,61:68,  141:144,181:182,60,140,180),
  c(3:4,  25:28,69:76,  145:148,183:184,24,68, 144),
  c(5:6,  29:32,77:84,  149:152,185:186,28,76, 148),
  c(7:8,  33:36,85:92,  153:156,187:188,32,84, 152),
  c(9:10, 37:40,93:100, 157:160,189:190,36,92, 156),
  c(11:12,41:44,101:108,161:164,191:192,40,100,160),
  c(13:14,45:48,109:116,165:168,193:194,44,108,164),
  c(15:16,49:52,117:124,169:172,195:196,48,116,168),
  c(17:18,53:56,125:132,173:176,197:198,52,124,172),
  c(19:20,57:60,133:140,177:180,199:200,56,132,176)
)
indirectSet4 = list(
  c(1:2,  21:24,61:68,  141:144,181:182,20,200),
  c(3:4,  25:28,69:76,  145:148,183:184,2, 182),
  c(5:6,  29:32,77:84,  149:152,185:186,4, 184),
  c(7:8,  33:36,85:92,  153:156,187:188,6, 186),
  c(9:10, 37:40,93:100, 157:160,189:190,8, 188),
  c(11:12,41:44,101:108,161:164,191:192,10,190),
  c(13:14,45:48,109:116,165:168,193:194,12,192),
  c(15:16,49:52,117:124,169:172,195:196,14,194),
  c(17:18,53:56,125:132,173:176,197:198,16,196),
  c(19:20,57:60,133:140,177:180,199:200,18,198)
)
indirectSet5 = list(
  c(1:2,  21:24,61:68,  141:144,181:182,140),
  c(3:4,  25:28,69:76,  145:148,183:184,68),
  c(5:6,  29:32,77:84,  149:152,185:186,76),
  c(7:8,  33:36,85:92,  153:156,187:188,84),
  c(9:10, 37:40,93:100, 157:160,189:190,92),
  c(11:12,41:44,101:108,161:164,191:192,100),
  c(13:14,45:48,109:116,165:168,193:194,108),
  c(15:16,49:52,117:124,169:172,195:196,116),
  c(17:18,53:56,125:132,173:176,197:198,124),
  c(19:20,57:60,133:140,177:180,199:200,132)
)
sets=list(directSet1, directSet2, directSet3, directSet4, directSet5,
          indirectSet1, indirectSet2, indirectSet3, indirectSet4, indirectSet5)
rm(directSet1, directSet2, directSet3, directSet4, directSet5,
   indirectSet1, indirectSet2, indirectSet3, indirectSet4, indirectSet5)
```


# Starting the simulation

For all models the person parameter theta needs to be set but unless data for a special sample is desired, a normal distribution should be used. The raw data are generated only once for all 10 anchor variants in order to avoid sample bias.

```{r include = FALSE}
work = data.frame()
for (i in 1:k) {
  theta = rnorm(N)
  sim.data <- simIrt(theta = theta, params = params, mod = "brm")
  #Combine item responses for the two groups into one file.
  sim.data.resp <- sim.data$resp
  work = rbind(work, cbind(sim.data.resp, i))
}
colnames(work) = c(paste("Item", 1:200, sep = ""), "i")
rm(theta, sim.data, sim.data.resp, params)
```

## Subsetting the data

In order to estimate how many items are needed for an anchoring approach, in the next step it will be determined which items are solved by which subsample.

```{r}
for (i in 1:length(sets)) { #for each anchor variant
  currentSet = unlist(sets[i], recursive=F) #extract the testforms
  data = work # make a copy of the data set
  for (j in 1:length(currentSet)) { #for each testform 
    currentTest = unlist(currentSet[j], recursive = F) #extract a vector of items from the testform
    data[, 1:200][(N * (j - 1) + 1):(N * j),-currentTest] = NA #set all but the extracted items to NA for a subsample
  }
  if(exists("allSamples")) allSamples = rbind(allSamples, data.frame(data, anchorType = i)) #in the first iteration create the result dataFrame
  else  allSamples = data.frame(data, anchorType = i) #in the following iterations append
}
rm(currentSet, currentTest, data, work)
```

# Running the test equating

```{r}
resultData = data.frame(realDiff=c(rep(-1.34, 20), rep(-0.36, 40), rep(-0.13, 80), rep(1.06, 40), rep(1.87, 20)))
system.time(
for (i in 1:length(sets)) {
  sample = allSamples[allSamples$anchorType == i,]
  subsamples = split(sample, sample$i)
  subsamples = lapply(subsamples, cleanData)
  if(i < 6) {res = directGetDiff(subsamples)}
  else {res = indirectGetDiff(subsamples,(31 - i))}
  resultData = cbind(resultData, res)
}
)
resultData$errorAT1 = (resultData[, 3] - resultData[, 1]) ^ 2
resultData$errorAT2 = (resultData[, 5] - resultData[, 1]) ^ 2
resultData$errorAT3 = (resultData[, 7] - resultData[, 1]) ^ 2
resultData$errorAT4 = (resultData[, 9] - resultData[, 1]) ^ 2
resultData$errorAT5 = (resultData[, 11] - resultData[, 1]) ^ 2
resultData$errorAT6 = (resultData[, 13] - resultData[, 1]) ^ 2
resultData$errorAT7 = (resultData[, 15] - resultData[, 1]) ^ 2
resultData$errorAT8 = (resultData[, 17] - resultData[, 1]) ^ 2
resultData$errorAT9 = (resultData[, 19] - resultData[, 1]) ^ 2
resultData$errorAT10 = (resultData[, 21] - resultData[, 1]) ^ 2
psych::describe(resultData[, (ncol(resultData) - 9):ncol(resultData)])
```

## Loading data of all iterations

```{r}
load("allResults.Rdata")
```

# Analyzing results of the test equating

## Boxplot for estimated Difficulties

```{r}
plotData = data.frame(type = c(rep("Direct", 50), rep("Indirect", 50)),
                      anchors = as.factor(rep(1:5, 20)),
                      diff = rnorm(100))
diffPlot = ggplot(plotData, aes(x = anchors, y = diff)) +
  geom_boxplot() + 
  scale_y_continuous(name = "Estimated average item difficulty")+
  scale_x_discrete(name = "Number of anchor items per test sets")+
  ggtitle("Boxplot for estimated Difficulties") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        text = element_text(size = 12),
        axis.title = element_text(face = "bold"),
        axis.text.x = element_text(size = 11)) +
  facet_grid(. ~ type)
diffPlot
```


## Creating a 2x5x5 matrix with the 3 factors ET, EV and RA

ET = Equating Type, EV = Equating variant, RA = Number of Rules (Item Difficulty)

```{r}
Iterations = 1000
estimates = data.frame(matrix(nrow=1000, ncol = 51))
estimates[,1] = 1:1000
colnames(estimates)[1] = "Iteration"
tmpString = ""
for (i in 1:2) {
  for (j in 1:5) {
    for (k in 1:5) {
      if(i == 1) tmpString = paste(paste("d", j, sep = ""), paste("v", k, sep = ""), sep = "_")
      else tmpString = paste(paste("i", j, sep = ""), paste("v", k, sep = ""), sep = "_")
      col = (i - 1) * 25 + (j - 1) * 5 + k + 1
      colnames(estimates)[col] = tmpString
    }
  }
}
rm(i, j, k, col, tmpString)
itemSets = list(1:20, 21:60, 61:140, 141:180, 181:200)
for (it in 1:Iterations) {
  for (set in 1:10) {
    for (rules in 1:5) {
      col = (set - 1) * 5 + rules + 1
      estimates[it, col] = mean(allResults[[it]][itemSets[[rules]], (set + 1)])
    }
  }
}
rm(Iterations, col, it, rules, set, itemSets)
```

## Splitting and trimming data (inner 95 %)

```{r}
d1_v1 = subset(estimates[2])
d1_v2 = subset(estimates[3])
d1_v3 = subset(estimates[4])
d1_v4 = subset(estimates[5])
d1_v5 = subset(estimates[6])
d2_v1 = subset(estimates[7])
d2_v2 = subset(estimates[8])
d2_v3 = subset(estimates[9])
d2_v4 = subset(estimates[10])
d2_v5 = subset(estimates[11])
d3_v1 = subset(estimates[12])
d3_v2 = subset(estimates[13])
d3_v3 = subset(estimates[14])
d3_v4 = subset(estimates[15])
d3_v5 = subset(estimates[16])
d4_v1 = subset(estimates[17])
d4_v2 = subset(estimates[18])
d4_v3 = subset(estimates[19])
d4_v4 = subset(estimates[20])
d4_v5 = subset(estimates[21])
d5_v1 = subset(estimates[22])
d5_v2 = subset(estimates[23])
d5_v3 = subset(estimates[24])
d5_v4 = subset(estimates[25])
d5_v5 = subset(estimates[26])
i1_v1 = subset(estimates[27])
i1_v2 = subset(estimates[28])
i1_v3 = subset(estimates[29])
i1_v4 = subset(estimates[30])
i1_v5 = subset(estimates[31])
i2_v1 = subset(estimates[32])
i2_v2 = subset(estimates[33])
i2_v3 = subset(estimates[34])
i2_v4 = subset(estimates[35])
i2_v5 = subset(estimates[36])
i3_v1 = subset(estimates[37])
i3_v2 = subset(estimates[38])
i3_v3 = subset(estimates[39])
i3_v4 = subset(estimates[40])
i3_v5 = subset(estimates[41])
i4_v1 = subset(estimates[42])
i4_v2 = subset(estimates[43])
i4_v3 = subset(estimates[44])
i4_v4 = subset(estimates[45])
i4_v5 = subset(estimates[46])
i5_v1 = subset(estimates[47])
i5_v2 = subset(estimates[48])
i5_v3 = subset(estimates[49])
i5_v4 = subset(estimates[50])
i5_v5 = subset(estimates[51])

highCutoff = 0.975
lowCutoff = 1 - highCutoff

d1_v1 = d1_v1 %>% filter(between(d1_v1, quantile(d1_v1, lowCutoff), quantile(d1_v1, highCutoff)))
d1_v2 = d1_v2 %>% filter(between(d1_v2, quantile(d1_v2, lowCutoff), quantile(d1_v2, highCutoff)))
d1_v3 = d1_v3 %>% filter(between(d1_v3, quantile(d1_v3, lowCutoff), quantile(d1_v3, highCutoff)))
d1_v4 = d1_v4 %>% filter(between(d1_v4, quantile(d1_v4, lowCutoff), quantile(d1_v4, highCutoff)))
d1_v5 = d1_v5 %>% filter(between(d1_v5, quantile(d1_v5, lowCutoff), quantile(d1_v5, highCutoff)))
d2_v1 = d2_v1 %>% filter(between(d2_v1, quantile(d2_v1, lowCutoff), quantile(d2_v1, highCutoff)))
d2_v2 = d2_v2 %>% filter(between(d2_v2, quantile(d2_v2, lowCutoff), quantile(d2_v2, highCutoff)))
d2_v3 = d2_v3 %>% filter(between(d2_v3, quantile(d2_v3, lowCutoff), quantile(d2_v3, highCutoff)))
d2_v4 = d2_v4 %>% filter(between(d2_v4, quantile(d2_v4, lowCutoff), quantile(d2_v4, highCutoff)))
d2_v5 = d2_v5 %>% filter(between(d2_v5, quantile(d2_v5, lowCutoff), quantile(d2_v5, highCutoff)))
d3_v1 = d3_v1 %>% filter(between(d3_v1, quantile(d3_v1, lowCutoff), quantile(d3_v1, highCutoff)))
d3_v2 = d3_v2 %>% filter(between(d3_v2, quantile(d3_v2, lowCutoff), quantile(d3_v2, highCutoff)))
d3_v3 = d3_v3 %>% filter(between(d3_v3, quantile(d3_v3, lowCutoff), quantile(d3_v3, highCutoff)))
d3_v4 = d3_v4 %>% filter(between(d3_v4, quantile(d3_v4, lowCutoff), quantile(d3_v4, highCutoff)))
d3_v5 = d3_v5 %>% filter(between(d3_v5, quantile(d3_v5, lowCutoff), quantile(d3_v5, highCutoff)))
d4_v1 = d4_v1 %>% filter(between(d4_v1, quantile(d4_v1, lowCutoff), quantile(d4_v1, highCutoff)))
d4_v2 = d4_v2 %>% filter(between(d4_v2, quantile(d4_v2, lowCutoff), quantile(d4_v2, highCutoff)))
d4_v3 = d4_v3 %>% filter(between(d4_v3, quantile(d4_v3, lowCutoff), quantile(d4_v3, highCutoff)))
d4_v4 = d4_v4 %>% filter(between(d4_v4, quantile(d4_v4, lowCutoff), quantile(d4_v4, highCutoff)))
d4_v5 = d4_v5 %>% filter(between(d4_v5, quantile(d4_v5, lowCutoff), quantile(d4_v5, highCutoff)))
d5_v1 = d5_v1 %>% filter(between(d5_v1, quantile(d5_v1, lowCutoff), quantile(d5_v1, highCutoff)))
d5_v2 = d5_v2 %>% filter(between(d5_v2, quantile(d5_v2, lowCutoff), quantile(d5_v2, highCutoff)))
d5_v3 = d5_v3 %>% filter(between(d5_v3, quantile(d5_v3, lowCutoff), quantile(d5_v3, highCutoff)))
d5_v4 = d5_v4 %>% filter(between(d5_v4, quantile(d5_v4, lowCutoff), quantile(d5_v4, highCutoff)))
d5_v5 = d5_v5 %>% filter(between(d5_v5, quantile(d5_v5, lowCutoff), quantile(d5_v5, highCutoff)))
i1_v1 = i1_v1 %>% filter(between(i1_v1, quantile(i1_v1, lowCutoff), quantile(i1_v1, highCutoff)))
i1_v2 = i1_v2 %>% filter(between(i1_v2, quantile(i1_v2, lowCutoff), quantile(i1_v2, highCutoff)))
i1_v3 = i1_v3 %>% filter(between(i1_v3, quantile(i1_v3, lowCutoff), quantile(i1_v3, highCutoff)))
i1_v4 = i1_v4 %>% filter(between(i1_v4, quantile(i1_v4, lowCutoff), quantile(i1_v4, highCutoff)))
i1_v5 = i1_v5 %>% filter(between(i1_v5, quantile(i1_v5, lowCutoff), quantile(i1_v5, highCutoff)))
i2_v1 = i2_v1 %>% filter(between(i2_v1, quantile(i2_v1, lowCutoff), quantile(i2_v1, highCutoff)))
i2_v2 = i2_v2 %>% filter(between(i2_v2, quantile(i2_v2, lowCutoff), quantile(i2_v2, highCutoff)))
i2_v3 = i2_v3 %>% filter(between(i2_v3, quantile(i2_v3, lowCutoff), quantile(i2_v3, highCutoff)))
i2_v4 = i2_v4 %>% filter(between(i2_v4, quantile(i2_v4, lowCutoff), quantile(i2_v4, highCutoff)))
i2_v5 = i2_v5 %>% filter(between(i2_v5, quantile(i2_v5, lowCutoff), quantile(i2_v5, highCutoff)))
i3_v1 = i3_v1 %>% filter(between(i3_v1, quantile(i3_v1, lowCutoff), quantile(i3_v1, highCutoff)))
i3_v2 = i3_v2 %>% filter(between(i3_v2, quantile(i3_v2, lowCutoff), quantile(i3_v2, highCutoff)))
i3_v3 = i3_v3 %>% filter(between(i3_v3, quantile(i3_v3, lowCutoff), quantile(i3_v3, highCutoff)))
i3_v4 = i3_v4 %>% filter(between(i3_v4, quantile(i3_v4, lowCutoff), quantile(i3_v4, highCutoff)))
i3_v5 = i3_v5 %>% filter(between(i3_v5, quantile(i3_v5, lowCutoff), quantile(i3_v5, highCutoff)))
i4_v1 = i4_v1 %>% filter(between(i4_v1, quantile(i4_v1, lowCutoff), quantile(i4_v1, highCutoff)))
i4_v2 = i4_v2 %>% filter(between(i4_v2, quantile(i4_v2, lowCutoff), quantile(i4_v2, highCutoff)))
i4_v3 = i4_v3 %>% filter(between(i4_v3, quantile(i4_v3, lowCutoff), quantile(i4_v3, highCutoff)))
i4_v4 = i4_v4 %>% filter(between(i4_v4, quantile(i4_v4, lowCutoff), quantile(i4_v4, highCutoff)))
i4_v5 = i4_v5 %>% filter(between(i4_v5, quantile(i4_v5, lowCutoff), quantile(i4_v5, highCutoff)))
i5_v1 = i5_v1 %>% filter(between(i5_v1, quantile(i5_v1, lowCutoff), quantile(i5_v1, highCutoff)))
i5_v2 = i5_v2 %>% filter(between(i5_v2, quantile(i5_v2, lowCutoff), quantile(i5_v2, highCutoff)))
i5_v3 = i5_v3 %>% filter(between(i5_v3, quantile(i5_v3, lowCutoff), quantile(i5_v3, highCutoff)))
i5_v4 = i5_v4 %>% filter(between(i5_v4, quantile(i5_v4, lowCutoff), quantile(i5_v4, highCutoff)))
i5_v5 = i5_v5 %>% filter(between(i5_v5, quantile(i5_v5, lowCutoff), quantile(i5_v5, highCutoff)))
```

## Merging and transforming data frame

```{r}
estimatesTrimmed = data.frame(d1_v1, d1_v2, d1_v3, d1_v4, d1_v5, d2_v1, d2_v2, d2_v3, d2_v4, d2_v5, d3_v1, d3_v2, d3_v3, d3_v4, d3_v5, d4_v1, d4_v2, d4_v3, d4_v4, d4_v5, d5_v1, d5_v2, d5_v3, d5_v4, d5_v5, i1_v1, i1_v2, i1_v3, i1_v4, i1_v5, i2_v1, i2_v2, i2_v3, i2_v4, i2_v5, i3_v1, i3_v2, i3_v3, i3_v4, i3_v5, i4_v1, i4_v2, i4_v3, i4_v4, i4_v5, i5_v1, i5_v2, i5_v3, i5_v4, i5_v5)

oneColTrimmed <- gather(estimatesTrimmed, key = "key", value = "value")

dfTrimmed = data.frame(type = c(rep(1, 23750), rep(2, 23750)), var = rep(c(rep(5, 4750), rep(4, 4750), rep(3, 4750), rep(2, 4750), rep(1, 4750)), 2), rule = rep(c(rep(1, 950), rep(2, 950), rep(3, 950), rep(4, 950), rep(5, 950)),50), oneColTrimmed)
```

## Descriptive statistics

d = direct equating, i = indirect equating, v = number of rules

```{r}
describe(estimatesTrimmed[, 1:50])

type1 = subset(estimatesTrimmed[1:25])
type1_oc = gather(type1, key = "key", value = "value")
describe(type1_oc)
mean(type1_oc$value)

type2 = subset(estimatesTrimmed[26:50])
type2_oc = gather(type2, key = "key", value = "value")
describe(type2_oc)
mean(type2_oc$value)

var1 = subset(estimatesTrimmed[c(1:5, 26:30)])
var1_oc = gather(var1, key = "key", value = "value")
describe(var1_oc)
mean(var1_oc$value)

var2 = subset(estimatesTrimmed[c(6:10, 31:35)])
var2_oc = gather(var2, key = "key", value = "value")
describe(var2_oc)
mean(var2_oc$value)

var3 = subset(estimatesTrimmed[c(11:15, 36:40)])
var3_oc = gather(var3, key = "key", value = "value")
describe(var3_oc)
mean(var3_oc$value)

var4 = subset(estimatesTrimmed[c(16:20, 41:45)])
var4_oc = gather(var4, key = "key", value = "value")
describe(var4_oc)
mean(var4_oc$value)

var5 = subset(estimatesTrimmed[c(21:25, 46:50)])
var5_oc = gather(var5, key = "key", value = "value")
describe(var5_oc)
mean(var5_oc$value)

rule1 = subset(estimatesTrimmed[c(1, 6, 11, 16, 21, 26, 31, 36, 41, 46)])
rule1_oc = gather(rule1, key = "key", value = "value")
describe(rule1_oc)
mean(rule1_oc$value)

rule2 = subset(estimatesTrimmed[c(2, 7, 12, 17, 22, 27, 32, 37, 42, 47)])
rule2_oc = gather(rule2, key = "key", value = "value")
describe(rule2_oc)
mean(rule2_oc$value)

rule3 = subset(estimatesTrimmed[c(3, 8, 13, 18, 23, 28, 33, 38, 43, 48)])
rule3_oc = gather(rule3, key = "key", value = "value")
describe(rule3_oc)
mean(rule3_oc$value)

rule4 = subset(estimatesTrimmed[c(4, 9, 14, 19, 24, 29, 34, 39, 44, 49)])
rule4_oc = gather(rule4, key = "key", value = "value")
describe(rule4_oc)
mean(rule4_oc$value)

rule5 = subset(estimatesTrimmed[c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50)])
rule5_oc = gather(rule5, key = "key", value = "value")
describe(rule5_oc)
mean(rule5_oc$value)
```

## ANOVA and contrasts

```{r}
# Transforming type, var and rule into factors

dfTrimmed$type = as.factor(dfTrimmed$type)
dfTrimmed$var = as.factor(dfTrimmed$var)
dfTrimmed$rule = as.factor(dfTrimmed$rule)

# Defining contrasts

c1 = c(1.5, 1.5, -1, -1, -1) # group contrast
c2 = c(2, -1, -2, -1, 2) # quadratic contrast

mat1 = cbind(c1)
mat2 = cbind(c2)

contrasts(dfTrimmed$var) = mat1
contrasts(dfTrimmed$rule) = mat2

# Conducting ANOVA

options(scipen = 999)
mod.new = aov(value ~ type * var * rule, data = dfTrimmed)
(sum.new = summary.aov(mod.new, split=list(var = list("Group_Comparison" = 1), rule = list("Quadratic" = 1))))
omega_squared(mod.new)

# Creating a table of the statistics from the ANOVA with more digits

Factor = c("ET", "EV", "EV_contrast", "RA", "RA_trend", "ET x EV", "-", "ET x RA", "-", "EV x RA", "-", "ET x EV x RA", "-", "Residual")
Degrees = round(as.numeric(sum.new[[1]]$Df), digits = 10)
Sum_Square = round(as.numeric(sum.new[[1]]$`Sum Sq`), digits = 10)
Mean_Square = round(as.numeric(sum.new[[1]]$`Mean Sq`), digits = 10)
F_Value = round(as.numeric(sum.new[[1]]$`F value`), digits = 10)
p_Value = round(as.numeric(sum.new[[1]]$`Pr(>F)`), digits = 10)
(df_ANOVA = data.frame(Factor, Degrees, Sum_Square, Mean_Square, F_Value, p_Value))

# Calculating contrasts

jmv::ANOVA(formula = value ~ type * var * rule, data = dfTrimmed, effectSize = "omega", postHoc = ~ type * var * rule, contrasts = list(list(var = "rule", type = "polynomial"), list(var = "var", type = "polynomial")))

TukeyHSD(mod.new, "var")

# Calculating the mean differences of the contrasts

(x1 = (mean(var1_oc$value) + mean(var2_oc$value) + mean(var3_oc$value))/3)
(x2 = (mean(var4_oc$value) + mean(var5_oc$value))/2)
```

# Effect sizes (Omega squared)

```{r}
ss_total = 226.5285716219 + 1344.3094846830 + 637.7495313988 + 106.9274055542 + 8.0231609941 + 20.1867696174 + 12.5088455094 + 8752.6126015211

omega.full.SS(dfm = 1, dfe = 237450,
msm = 226.5285716219, mse = 0.0368608659,
sst = (ss_total), a = .05) # ET

omega.full.SS(dfm = 4, dfe = 237450,
msm = 336.0773711708, mse = 0.0368608659,
sst = (ss_total), a = .05) # EV

omega.full.SS(dfm = 1, dfe = 237450,
msm = 1225.9539917622, mse = 0.0368608659,
sst = (ss_total), a = .05) # EV Contrast

omega.full.SS(dfm = 4, dfe = 237450,
msm = 159.4373828497, mse = 0.0368608659,
sst = (ss_total), a = .05) # RA

omega.full.SS(dfm = 1, dfe = 237450,
msm = 450.5662092836, mse = 0.0368608659,
sst = (ss_total), a = .05) # RA Trend

omega.full.SS(dfm = 4, dfe = 237450,
msm = 26.7318513886, mse = 0.0368608659,
sst = (ss_total), a = .05) # ET x EV

omega.full.SS(dfm = 4, dfe = 237450,
msm = 2.0057902485, mse = 0.0368608659,
sst = (ss_total), a = .05) # ET x RA

omega.full.SS(dfm = 16, dfe = 237450,
msm = 1.2616731011, mse = 0.0368608659,
sst = (ss_total), a = .05) # EV x RA

omega.full.SS(dfm = 16, dfe = 237450,
msm = 0.7818028443, mse = 0.0368608659,
sst = (ss_total), a = .05) # ET x EV x RA

omega.full.SS(dfm = 1, dfe = 237450,
msm = 16.2^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v5v4

omega.full.SS(dfm = 1, dfe = 237450,
msm = 29.9^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v5v3

omega.full.SS(dfm = 1, dfe = 237450,
msm = 109.1^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v5v2

omega.full.SS(dfm = 1, dfe = 237450,
msm = 157.2^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v5v1

omega.full.SS(dfm = 1, dfe = 237450,
msm = 13.7^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v4v3

omega.full.SS(dfm = 1, dfe = 237450,
msm = 92.8^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v4v2

omega.full.SS(dfm = 1, dfe = 237450,
msm = 140.9^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v4v1

omega.full.SS(dfm = 1, dfe = 237450,
msm = 79.1^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v3v2

omega.full.SS(dfm = 1, dfe = 237450,
msm = 127.2^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v3v1

omega.full.SS(dfm = 1, dfe = 237450,
msm = 48.1^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # v2v1

omega.full.SS(dfm = 1, dfe = 237450,
msm = 48.5894^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r1r2

omega.full.SS(dfm = 1, dfe = 237450,
msm = 48.5771^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r1r3

omega.full.SS(dfm = 1, dfe = 237450,
msm = 18.4023^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r1r4

omega.full.SS(dfm = 1, dfe = 237450,
msm = 64.1836^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r1r5

omega.full.SS(dfm = 1, dfe = 237450,
msm = 0.0123^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r2r3

omega.full.SS(dfm = 1, dfe = 237450,
msm = 30.1872^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r2r4

omega.full.SS(dfm = 1, dfe = 237450,
msm = 112.7731^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r2r5

omega.full.SS(dfm = 1, dfe = 237450,
msm = 30.1479^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r3r4

omega.full.SS(dfm = 1, dfe = 237450,
msm = 112.7608^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r3r5

omega.full.SS(dfm = 1, dfe = 237450,
msm = 82.5859^2 * 0.0368608659, mse = 0.0368608659,
sst = (ss_total), a = .05) # r4r5
```

# Plotting data

```{r}
ggplot(dfTrimmed, aes(x = type, y = value, fill = type))  +
  ylim(0, 2.5) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_line(colour = "grey")) +
  geom_boxplot() +
  scale_x_discrete(labels = c("direkt", "indirekt")) +
  scale_y_continuous(breaks=c(0.5, 1, 1.5, 2, 2.5), limits = c(0,2.5)) +
  scale_fill_manual(values = c(rep("lightgrey", 2))) +
  labs(x = "Equating-Typ", y = "Abweichungsquadrat") +
  guides(fill=FALSE)

ggplot(dfTrimmed, aes(x = var, y = value, fill = var))  +
  ylim(0, 2.5) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_line(colour = "grey")) +
  geom_boxplot() +
  scale_x_discrete() +
  scale_y_continuous(breaks=c(0.5, 1, 1.5, 2, 2.5), limits = c(0,2.5)) +
  scale_fill_manual(values = c(rep("lightgrey", 5))) +
  labs(x = "Equating-Variante", y = "Abweichungsquadrat") +
  guides(fill=FALSE)

ggplot(dfTrimmed, aes(x = rule, y = value, fill = rule))  +
  ylim(0, 2.5) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_line(colour = "grey")) +
  geom_boxplot() +
  scale_x_discrete() +
  scale_y_continuous(breaks=c(0.5, 1, 1.5, 2, 2.5), limits = c(0,2.5)) +
  scale_fill_manual(values = c(rep("lightgrey", 5))) +
  labs(x = "Regelanzahl", y = "Abweichungsquadrat") +
  guides(fill=FALSE)

ggplot(dfTrimmed, aes(x = var, y = value, fill = type))  +
  ylim(0, 2.5) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_line(colour = "grey"), legend.position = "top") +
  geom_boxplot() +
  scale_x_discrete() +
  scale_y_continuous(breaks=c(0.5, 1, 1.5, 2, 2.5), limits = c(0,2.5)) +
  scale_fill_manual(values = c("lightgrey", "#A4A4A4"), labels = c("direkt", "indirekt")) +
  labs(x = "Equating-Variante", y = "Abweichungsquadrat", fill = "Equating-Typ")

ggplot(dfTrimmed, aes(x = type, y = value, fill = var))  +
  ylim(0, 2.5) +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_line(colour = "grey"), legend.position = "top", legend.key.width = unit(0.25,"cm")) +
  geom_boxplot() +
  scale_x_discrete(labels = c("direkt", "indirekt")) +
  scale_y_continuous(breaks=c(0.5, 1, 1.5, 2, 2.5), limits = c(0,2.5)) +
  #scale_fill_manual(values = c("#A9BCF5", "#F5A9A9", "#F2F5A9", "#A9F5A9", "#F5D0A9")) +
  scale_fill_manual(values = c("#FFFFFF", "#E6E6E6", "#BDBDBD", "#A4A4A4", "#848484")) +
  labs(x = "Equating-Typ", y = "Abweichungsquadrat", fill = "Equating-Variante")
```